The Impact of Alternation

### 2017-05-09 draft.tex
ok

\begin{abstract}
Alternating automata have been widely used to model and verify systems
that handle data from finite domains, such as communication protocols
or hardware. The main advantage of the alternating model of
computation is that complementation is possible in linear time, thus
allowing to concisely encode trace inclusion problems that occur often
in verification. In this paper we consider alternating automata over
infinite alphabets, whose transition rules are formulae in a combined
theory of booleans and some infinite data domain, that relate past and
current values of the data variables. The data theory is not fixed,
but rather it is a parameter of the class. We show that union,
intersection and complementation are possible in linear time in this
model and, though the emptiness problem is undecidable, we provide two
efficient semi-algorithms, inspired by two state-of-the-art
abstraction refinement model checking methods: lazy predicate
abstraction and the IMPACT algorithm. We have implemented both methods
and report the results of an experimental comparison.
\end{abstract}

@@

A \emph{signature} $\sig = (\ssorts{\sig},\sfuns{\sig})$ consists of a
set $\ssorts{\sig}$ of \emph{sort symbols} and a set $\sfuns{\sig}$ of
sorted \emph{function symbols}. To simplify the presentation, we
assume w.l.o.g. that $\ssorts{\sig}$ consists of two distinct sorts
$\Data$ and $\Bool$ and each function symbol $f \in \sfuns{\sig}$ has
$\#(f) \geq 0$ arguments of sort $\Data$ and return value $\sigma(f)
\in \ssorts{\sig}$. If $\#(f)=0$ then $f$ is a \emph{constant}. We
consider the constants $\mathbf{0}$ of sort $\Data$ and $\top,\bot$ of
sort $\Bool$.

@@

In the rest of this section we fix an interpretation $\I$ and a finite
alphabet $\Sigma$ of \emph{input events}. Given a finite set $\vec{x}
\subset \vars$ of variables of sort $\Data$, let $\vec{x} \mapsto
\Data^\I$ be the set of valuations of the variables $\vec{x}$ and
$\Sigma[\vec{x}] = \Sigma \times (\vec{x} \mapsto \Data^\I)$ be the
set of \emph{data symbols}. A \emph{data word} (or, simply, a word) is
a finite sequence $(a_1,\nu_1)(a_2,\nu_2) \ldots (a_n,\nu_n)$ of data
symbols, where $a_1,\ldots,a_n \in \Sigma$ and $\nu_1,\ldots,\nu_n :
\vec{x} \rightarrow \Data^\I$ are valuations. We denote by
$\varepsilon$ the empty sequence, by $\Sigma^*$ the set of finite
sequences of input events and by $\Sigma[\vec{x}]^*$ the set of data
words over $\vec{x}$.

@@

An \emph{alternating data automaton} (ADA) or, simply, an automaton,
is a tuple $\A =
\tuple{\vec{x},Q,\iota,F,\Delta}$, where: \begin{compactitem}
%
\item $\vec{x} \subset \vars$ is a finite set of variables of sort
  $\Data$,
%
\item $Q$ is a finite set of variables of sort $\Bool$, called
  \emph{states},
%
\item $\iota \in \Form^+(Q,\emptyset)$ is positive boolean combination
  of states defining the \emph{initial configurations},
%
\item $F \subseteq Q$ is a set of \emph{final states}, and
%
\item $\Delta : Q \times \Sigma \rightarrow
  \Form^+(Q,\overline{\vec{x}}\cup\vec{x})$ is a \emph{transition
    function}, where $\overline{\vec{x}}$ is the set $\{\overline{x}
  \mid x \in \vec{x}\}$.
\end{compactitem}
Intuitively, in each formula $\Delta(q,a)$ describing a transition
rule, the variables $\overline{\vec{x}}$ track the previous and
$\vec{x}$ the current values of the variables of $\A$. Observe
further, that the initial values of the variables are left
unconstrained, as the initial configuration does not contain free
data variables. The size of $\A$ is defined as $\len{\A} = \len{\iota}
+ \sum_{(q,a) \in Q\times\Sigma}\len{\Delta(q,a)}$.

@@

Consequently, we give up on the guarantee for termination and develop
semi-algorithms that are supposed to be: \begin{compactenum}
\item \emph{sound}: given an automaton $\A$, if $L(\A) \neq\emptyset$,
  the procedure will terminate and return a word $w \in L(\A)$, and
%
\item \emph{complete relative to termination}: if the procedure terminates
  without returning such a word, then $L(\A) = \emptyset$. 
\end{compactenum}

@@

The solution we adopt is abstraction to ensure the termination of
invariant computations. However, it is worth pointing out from the
start that abstraction will allow us to build invariants which are not
necessarily safety invariants. To meet the latter condition, we resort
to counterexample guided abstraction refinement (CEGAR).

@@

In general, given an unsatisfiable conjunction $\pi \equiv
\phi_1(X_0,X_1) \wedge \phi_2(X_1,X_2) \wedge \ldots \wedge
\phi_n(X_{n-1},X_n)$ of time-stamped variables $X_i = \set{x_i \mid x
  \in X}$ of any sort, a solution of the \emph{interpolation problem}
$\pi$, simply called an \emph{interpolant}, is a tuple $\tuple{I_0(X),
  I_1(X), \ldots, I_n(X)}$ such that:\begin{inparaenum}[(i)]
\item $I_0 \equiv \top$,  
%
\item $I_{i-1}[X_{i-1}/X] \wedge \phi_i(X_{i-1},X_i) \models
  I_i[X_i/X]$, for all $i \in [1,n]$, and
%
\item $I_n \equiv \bot$. 
\end{inparaenum}
A classical method for abstraction refinement is to add the elements
of the interpolant obtained from a proof of spuriousness to the set of
predicates. This guarantees progress, meaning that the particular
spurious counterexample, from which the interpolant was generated,
will never be revisited in the future. Though not always, in many
practical test cases this progress property eventually yields a safety
invariant.

@@

{\scriptsize\begin{algorithmic}[1]
  \State $\preds \leftarrow \set{\bot}$, $\worklist \leftarrow
  \tuple{\rootNode}$, $\Art = \tuple{N,E,\Lambda,\lhd}$, $N \leftarrow
  \emptyset$, $E \leftarrow \emptyset$, $\Lambda \leftarrow
  \set{(\rootNode,\iota)}$, $\lhd \leftarrow \emptyset$

@@

  \If{$\Lambda(n) \wedge \bigwedge_{q \in Q \setminus F} (q
    \rightarrow \bot)$ is satisfiable} \Comment{counterexample found}

@@

  \State $p \leftarrow \Call{FindPivot}{u,\Art}$

@@

  \State $\preds \leftarrow \preds \cup \set{I_1,\ldots,I_\ell}$,
  where $\tuple{\top,I_1,\ldots,I_\ell,\bot}$ is the interpolant for
  $\overline{\pi}(v)$

@@

  \For{$a \in \Sigma$}\label{ln:forall-events}

@@

Formally, an ART $\Art = \tuple{N,E,\Lambda,\lhd}$ is a tree, where
$N$ is a set of nodes, $E \subseteq N \times \Sigma \times N$ is a set
of edges, $\Lambda : N \rightarrow \Form^+(Q,\vec{x})$ is a labeling
of the nodes with formulae and $\lhd \subseteq N \times N$ is a set
of covering edges. The root of the tree is labeled with the initial
configuration $\iota$ and each node $n \in N$ corresponds to a unique
path from the root to $n$, labeled by a sequence $u$ of input
events. The least unfeasible suffix of $u$ is the smallest sequence $v
= a_1 \ldots a_k$, such that $u = wv$, for some $w \in \Sigma^*$ and
the formula (\ref{eq:pivot}) below is unsatisfiable. The \emph{pivot}
of $u$ is the node $p$ corresponding to the start of $v$ in $u$.
\begin{equation}\label{eq:pivot}
  \overline{\pi}(v) \equiv \Lambda(p)[Q_0/Q] \wedge
  \theta_1(Q_0 \cup Q_1,\vec{x}_0 \cup \vec{x}_1) \wedge \ldots \wedge
  \theta_{k+1}(Q_k)
\end{equation}
where $\theta_1,\ldots,\theta_{k+1}$ are defined as in
(\ref{eq:interpolation-problem}), for $\theta_0 \equiv
\Lambda(p)[Q_0/Q]$.

With these considerations, Algorithm \ref{alg:predabs} uses a worklist
iteration to build an ART. We keep newly expanded nodes of $\Art$ in a
use a queue $\worklist$ to ensure a breadth-first exploration
strategy, which guarantees that the shortest counterexamples are
explored first. When the search encounters a counterexample $u$, it is
checked for spuriousness. If the counterexample is real, the procedure
returns a data word $w \in L(\A)$, which interleaves the input events
of $u$ with the data valuations from the model of $\Accept{\A}(u)$
(since $u$ is feasible, clearly $\Accept{\A}(u)$ is
satisfiable). Otherwise, if $u$ is spurious, we compute its pivot $p$,
add the interpolants for the least unfeasible suffix of $u$ to the set
of predicates $\preds$, remove and recompute the subtree of $\Art$
rooted at $p$.

Termination of Algorithm \ref{alg:predabs} depends on the ability of a
given interpolating decision procedure for the combined boolean and
data theory $\theory(\sig,\I)$ to provide interpolants that yield a
safety invariant, provided that $L(\A) = \emptyset$. In this case, we
use the covering relation $\lhd$ to ensure that, when a newly
generated node is covered by a node already in $N$, it is not added to
the worklist, thus cutting the current branch of the search. Formally,
for any two nodes $n,m \in N$, we have $n \lhd m$ iff
$\AbsPost{\A}(\Lambda(n),a) \models \Lambda(m)$ for some $a \in
\Sigma$, in other words, if $n$ has a successor whose label entails
the label of $m$. The soundness and relative completness of Algorithm
\ref{alg:predabs} are proved by the following theorem:

@@

>>>>

\begin{abstract}
Alternating automata have been widely used to model and verify systems
that handle data from finite domains, such as communication protocols
or hardware. The main advantage of the alternating model of
computation is that complementation is possible in linear time, thus
allowing to concisely encode trace inclusion problems that occur often
in verification. In this paper we consider alternating automata over
infinite alphabets, whose transition rules are formulae in a combined
theory of booleans and some infinite data domain, that relate past and
current values of the data variables. The data theory is not fixed,
but rather it is a parameter of the class. We show that union,
intersection and complementation are possible in linear time in this
model and, though the emptiness problem is undecidable, we provide two
efficient semi-algorithms, inspired by two state-of-the-art
abstraction refinement model checking methods: lazy predicate
abstraction \cite{HJMS02} and the \impact~ semi-algorithm
\cite{mcmillan06}. We have implemented both methods and report the
results of an experimental comparison.
\end{abstract}

@@

A \emph{signature} $\sig = (\ssorts{\sig},\sfuns{\sig})$ consists of a
set $\ssorts{\sig}$ of \emph{sort symbols} and a set $\sfuns{\sig}$ of
sorted \emph{function symbols}. To simplify the presentation, we
assume w.l.o.g. that $\ssorts{\sig}$ consists of two distinct sorts
$\Data$ and $\Bool$ and each function symbol $f \in \sfuns{\sig}$ has
$\#(f) \geq 0$ arguments of sort $\Data$ and return value $\sigma(f)
\in \ssorts{\sig}$. If $\#(f)=0$ then $f$ is a \emph{constant}. We
consider the constants $\top$ and $\bot$ of sort $\Bool$.

@@

In the rest of this section we fix an interpretation $\I$ and a finite
alphabet $\Sigma$ of \emph{input events}. Given a finite set $\vec{x}
\subset \vars$ of variables of sort $\Data$, let $\vec{x} \mapsto
\Data^\I$ be the set of valuations of the variables $\vec{x}$ and
$\Sigma[\vec{x}] = \Sigma \times (\vec{x} \mapsto \Data^\I)$ be the
set of \emph{data symbols}. A \emph{data word} (word in the sequel) is
a finite sequence $(a_1,\nu_1)(a_2,\nu_2) \ldots (a_n,\nu_n)$ of data
symbols, where $a_1,\ldots,a_n \in \Sigma$ and $\nu_1,\ldots,\nu_n :
\vec{x} \rightarrow \Data^\I$ are valuations. We denote by
$\varepsilon$ the empty sequence, by $\Sigma^*$ the set of finite
sequences of input events and by $\Sigma[\vec{x}]^*$ the set of data
words over $\vec{x}$.

@@

An \emph{alternating data automaton} (ADA or automaton in the sequel)
is a tuple $\A = \tuple{\vec{x},Q,\iota,F,\Delta}$,
where: \begin{compactitem}
%
\item $\vec{x} \subset \vars$ is a finite set of variables of sort
  $\Data$,
%
\item $Q$ is a finite set of variables of sort $\Bool$, called
  \emph{states},
%
\item $\iota \in \Form^+(Q,\emptyset)$ is positive boolean combination
  of states defining the \emph{initial configurations},
%
\item $F \subseteq Q$ is a set of \emph{final states}, and
%
\item $\Delta : Q \times \Sigma \rightarrow
  \Form^+(Q,\overline{\vec{x}}\cup\vec{x})$ is a \emph{transition
    function}, where $\overline{\vec{x}}$ is the set $\{\overline{x}
  \mid x \in \vec{x}\}$.
\end{compactitem}
Intuitively, in each formula $\Delta(q,a)$ describing a transition
rule, the variables $\overline{\vec{x}}$ track the previous and
$\vec{x}$ the current values of the variables of $\A$. Observe
further, that the initial values of the variables are left
unconstrained, as the initial configuration does not contain free
data variables. The size of $\A$ is defined as $\len{\A} = \len{\iota}
+ \sum_{(q,a) \in Q\times\Sigma}\len{\Delta(q,a)}$.

@@

Consequently, we give up on the guarantee for termination and build
semi-algorithms that meet the requirements below: \begin{compactenum}[(i)]
\item given an automaton $\A$, if $L(\A) \neq\emptyset$, the procedure
  will terminate and return a word $w \in L(\A)$, and
%
\item if the procedure terminates without returning such a word, then
  $L(\A) = \emptyset$.
\end{compactenum}

@@

The solution we adopt is abstraction to ensure the termination of
invariant computations. However, it is worth pointing out from the
start that abstraction alone will only allow us to build invariants
that are not necessarily safety invariants. To meet the latter
condition, we resort to counterexample guided abstraction refinement
(CEGAR).

@@

In general, given an unsatisfiable conjunction $\pi \equiv
\phi_1(X_0,X_1) \wedge \phi_2(X_1,X_2) \wedge \ldots \wedge
\phi_n(X_{n-1},X_n)$ of time-stamped variables $X_i = \set{x_i \mid x
  \in X}$ of any sort, a solution of the \emph{interpolation problem}
$\pi$, simply called an \emph{interpolant}, is a tuple $\tuple{I_0(X),
  I_1(X), \ldots, I_n(X)}$ such that:\begin{inparaenum}[(i)]
\item $I_0 \equiv \top$,  
%
\item $I_{i-1}[X_{i-1}/X] \wedge \phi_i(X_{i-1},X_i) \models
  I_i[X_i/X]$, for all $i \in [1,n]$, and
%
\item $I_n \equiv \bot$. 
\end{inparaenum}

\paragraph{\em Remark} 
We deviate from the standard notion of \emph{generalized interpolants}
\cite{mcmillan06}, which requires $\fv{}{I_i} \subseteq \fv{}{\phi_i}
\cap \fv{}{\phi_{i+1}}$, for each $i \in [1,n-1]$, and assume further
that the variables have been already renamed by removing the
time-stamps. \qed

A classical method for abstraction refinement is to add the elements
of the interpolant obtained from a proof of spuriousness to the set of
predicates. This guarantees progress, meaning that the particular
spurious counterexample, from which the interpolant was generated,
will never be revisited in the future. Though not always, in many
practical test cases this progress property eventually yields a safety
invariant.

@@

{\scriptsize\begin{algorithmic}[1]

  \State let $\Art = \tuple{N,E,\Lambda,\lhd}$ be an ART

  \State initially $N = E = \lhd = \emptyset$, $\Lambda =
  \set{(\rootNode,\iota)}$, $\preds = \set{\bot}$, $\worklist =
  \tuple{\rootNode}$,

@@

  \If{$\Lambda(n) \wedge \bigwedge_{q \in Q \setminus F} (q
    \rightarrow \bot)$ is satisfiable} \Comment{counterexample candidate}

@@

  \State $p \leftarrow \Call{FindPivot}{u,\Art}$
  \label{ln:pivot}

@@

  \State $\preds \leftarrow \preds \cup \set{I_1,\ldots,I_\ell}$,
  where $\tuple{\top,I_1,\ldots,I_\ell,\bot}$ is an interpolant for
  $\overline{\pi}(v)$

@@

  \For{$a \in \Sigma$}\label{ln:forall-events}
  \Comment{expand $n$}

@@

Formally, an ART $\Art = \tuple{N,E,\Lambda,\lhd}$ is a tree, where
$N$ is a set of nodes, $E \subseteq N \times \Sigma \times N$ is a set
of edges, $\Lambda : N \rightarrow \Form^+(Q,\vec{x})$ is a labeling
of the nodes with formulae and $\lhd \subseteq N \times N$ is a set of
covering edges. We require that $(N,E)$ is a directed tree, rooted at
$\rootNode \in N$, where $\Lambda(\rootNode) = \iota$. Then each node
$n \in N$ corresponds to a unique path from the root to $n$, labeled
by a sequence $u$ of input events. The least unfeasible suffix of $u$
is the smallest sequence $v = a_1 \ldots a_k$, such that $u = wv$, for
some $w \in \Sigma^*$ and the formula (\ref{eq:pivot}) below is
unsatisfiable. The \emph{pivot} of $u$ is the node $p$ corresponding
to the start of $v$ in $u$.
\begin{equation}\label{eq:pivot}
  \overline{\pi}(v) \equiv \Lambda(p)[Q_0/Q] \wedge
  \theta_1(Q_0 \cup Q_1,\vec{x}_0 \cup \vec{x}_1) \wedge \ldots \wedge
  \theta_{k+1}(Q_k)
\end{equation}
where $\theta_1,\ldots,\theta_{k+1}$ are defined as in
(\ref{eq:interpolation-problem}), for $\theta_0 \equiv
\Lambda(p)[Q_0/Q]$.

With these considerations, Algorithm \ref{alg:predabs} uses a worklist
iteration to build an ART. We keep newly expanded nodes of $\Art$ in a
use a queue $\worklist$ to ensure a breadth-first exploration
strategy, which guarantees that the shortest counterexamples are
explored first. When the search encounters a counterexample $u$, it is
checked for spuriousness. If the counterexample is real, the procedure
returns a data word $w \in L(\A)$, which interleaves the input events
of $u$ with the data valuations from the model of $\Accept{\A}(u)$
(since $u$ is feasible, clearly $\Accept{\A}(u)$ is
satisfiable). Otherwise, if $u$ is spurious, we compute its pivot $p$
(line \ref{ln:pivot}), add the interpolants for the least unfeasible
suffix of $u$ to the set of predicates $\preds$, remove and recompute
the subtree of $\Art$ rooted at $p$.

Termination of Algorithm \ref{alg:predabs} depends on the ability of a
given interpolating decision procedure for the combined boolean and
data theory $\theory(\sig,\I)$ to provide interpolants that yield a
safety invariant, provided that $L(\A) = \emptyset$. In this case, we
use the covering relation $\lhd$ to ensure that, when a newly
generated node is covered by a node already in $N$, it is not added to
the worklist, thus cutting the current branch of the search. Formally,
for any two nodes $n,m \in N$, we have $n \lhd m$ iff
$\AbsPost{\A}(\Lambda(n),a) \models \Lambda(m)$ for some $a \in
\Sigma$, in other words, if $n$ has a successor whose label entails
the label of $m$. The correctness criteria of Algorithm
\ref{alg:predabs} are showed in the following theorem:

@@

\section{Checking ADA Emptiness with \impact}

As pointed out by a number of authors (e.g.\ \cite{mcmillan06}) the
bottleneck of predicate abstraction is the high cost of existential
quantifier elimination required to compute the $\Post{\A}(\phi,a)$
function --- we recall that $\Post{\A}(\phi,a)$ is defined by the
formula $\overline{\vec{x}} ~.~
\Delta(\phi[\overline{\vec{x}}/\vec{x}],a)$. The main idea of the
\impact~ procedure is that interpolation can be used instead
quantifier elimination, because an interpolant over-approximates the
post-images along the counterexample path in a way that is sufficient
to prove its spuriousness.

With this in mind, Algorithm \ref{alg:impact} builds an ART for the
given automaton and uses interpolants, instead of abstract
post-images, to label its nodes. To speedup the computation we use two
additional labeling functions for bookkeeping. Formally, the ART
structure used by Algorithm \ref{alg:impact} is defined as $\Art =
\tuple{N,E,\Lambda,R,T,\lhd}$, where $N,E,\Lambda$ and $\lhd$ are
defined as for Algorithm \ref{alg:predabs} and,
moreover: \begin{compactitem}
\item $R : N \rightarrow 2^Q$ is a labeling of nodes with replacement
  sets, such that $R(\rootNode) = \fv{\Bool}{\iota}$, 
%
\item $T : E \rightarrow \bigcup_{i=0}^\infty
  \Form(Q_i,\vec{x}_i,Q_{i+1},\vec{x}_{i+1})$ is labeling of edges
  with time-stamped formulae, in which $Q_i$ may occur under negation,
  and
%
\item for each edge $e = (n,a,m) \in E$, if $n$ lies at
  depth\footnote{We consider that the root node has zero depth.}
  $k\geq0$ in the tree $(N,E)$, then we have:
  \[\begin{array}{rcl}
  T(e) & = & \Theta(n,k) \\
  R(m) & = & \bigcup_{q \in R(n)}\fv{\Bool}{\Delta(q,a)} \text{ where} \\
  \Theta(n,k) & \equiv & \bigwedge_{q \in R(n)} (q_k \rightarrow
  \Delta(q,a)[Q_{k+1}/Q,\vec{x}_{k}/\overline{\vec{x}},\vec{x}_{k+1}/\vec{x}])
  \end{array}\]
\end{compactitem}
For simplicity, we write $T(n,a,m)$ for $T(e)$, when $e$ is
$(n,a,m)$. The idea behind this additional bookkeeping is that, given
a sequence $u = a_1\ldots a_k \in \Sigma^*$ that labels the path
from the root to a node $n$ in $\Art$, the conjunction: 
\[\mu(n) \equiv \Lambda(\rootNode) \wedge T(\rootNode,a_1,n_1) \wedge \ldots
T(n_{k-1},a_k,n_k) \wedge \bigwedge_{q \in R(n_k) \setminus F}(q
\rightarrow \bot)\] is equivalent to the interpolation problem
$\pi(u)$. We assume furthermore a total alphabetical order $\prec$ on
$\Sigma$ and lift it to the total lexicographical order $\prec^*$ on
$\Sigma^*$. We write $n \prec^* m$ if the sequence labeling the path
from $\rootNode$ to $n$ is strictly smaller, in the lexicographical
order, than the sequence labeling the path from $\rootNode$ to $m$. A
node $n \in N$ is \emph{covered by $m$} if $(n,m) \in \lhd$ and simply
covered when $m$ is not important. A node $n$ is \emph{closed} if for
all nodes $m \in N$ such that $m \prec^* n$, either $\Lambda(n)
\not\models \Lambda(m)$ or $(n,m) \in \lhd$.

{\scriptsize\begin{algorithmic}[1]

  \State let $\Art = \tuple{N,E,\Lambda,R,T,\lhd}$ be an ART

  \State initially $N = E = T = \lhd = \emptyset$, $\Lambda =
  \set{(\rootNode,\iota)}$, $R = \fv{\Bool}{\iota[Q_0/Q]}$,
  $\worklist=\set{\rootNode}$

  \While{$\worklist \neq \emptyset$}

  \For{$m \in N$ such that $m \prec^* n$}
  \Comment{close $n$}

  \If{$\Lambda(n) \models \Lambda(m)$}

  \State $\lhd \leftarrow \lhd \setminus \set{ (p,q) \in \lhd \mid
    \text{$q$ is a successor of $n$}}$

  \State $\lhd \leftarrow \lhd \cup \set{(n,m)}$

  \If{$n$ is not covered}

  \If{$R(n) \cap F \neq \emptyset$} 
  \Comment{counterexample candidate}

  \If{$\mu(n)$ is satisfiable} 
  \Comment{counterexample is real}

  \State get model $\beta : Q_i \rightarrow \booli$, $\{\nu_i :
  \vec{x}_i \rightarrow \Data^\I \mid i=0\ldots k\}$ of $\mu(n)$

  \State {\bf return} $w = (a_1,\nu_1) \ldots (a_k,\nu_k)$
  \Comment{$w \in L(\A)$ by construction}

  \State let $\rootNode,n_1,\ldots,n_k$ be the sequence of nodes from $\rootNode$ to $n$

  \State let $\tuple{\top,I_1,\ldots,I_k,I_{k+1}\equiv\bot}$ be an interpolant for $\mu(n)$

  \For{$i=1,\ldots,k+1$}

  \If{$\Lambda(n_i) \not\models I_i$}

  \State $\lhd \leftarrow \lhd \setminus \set{(m,n_i) \in \lhd \mid m \in N}$
  \Comment{uncover $n_i$}

  \State $\Lambda(n_i) \leftarrow \Lambda(n_i) \wedge I_i$
  \Comment{strenghten the label of $n_i$}

  \For{$a \in \Sigma$} 
  \Comment{expand $n$}

  \State let $s$ be a fresh node and $e \leftarrow (n,a,s)$

  \State $E \leftarrow E \cup \set{e}$

  \State $\Lambda \leftarrow \Lambda \cup \set{(s,\top)}$

  \State $T \leftarrow T \cup \set{(e, \Theta(n,k))}$

  \State $R \leftarrow R \cup \{(s,\bigcup_{q \in R(n)}\fv{\Bool}{\Delta(q,a)})\}$

  \State enqueue $s$ into $\worklist$

  \EndWhile  
\end{algorithmic}}
\caption{\impact~ for ADA Emptiness}
\label{alg:impact}
\end{algorithm}
